% REMEMBER: You must not plagiarise anything in your report. Be extremely careful.

\documentclass{l4proj}

    
%
% put any additional packages here
%
\usepackage[l2tabu, orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage{listings-rust}
\usepackage{dirtree}
\usepackage{pgf}
\usepackage[bottom]{footmisc}
\lstset{language=Rust}

\begin{document}

%==============================================================================
%% METADATA
\title{The IETF Transport Service API}
\author{Stuart Reilly}
\date{\today}
\newcommand{\asyncawait}{\texttt{async/await}}
\newcommand{\preconnection}{\emph{Preconnection}}
\newcommand{\connection}{\emph{Connection}}
\newcommand{\listener}{\emph{Listener}}
\newcommand{\framer}{\emph{Framer}}
\newcommand{\Endpoint}{\emph{Endpoint}}
\newcommand{\encode}{\emph{Encode}}
\newcommand{\decode}{\emph{Decode}}

\maketitle

%==============================================================================
%% ABSTRACT
\begin{abstract}
    Network programming has been build around the Berkeley Sockets API since its inception in 1983.
    While the simplicity of this API, which abstracted file I/O and network I/O to a similar interface,
    is a testament to its longevity, modern network programming has changed in a manner which requires
    amounts of boilerplate code.
    This projects aims to investigate the possibility of a high-level API which removes the boilerplate
    associated with Berkeley sockets, and provide a protocol-agnostic, asynchronous API to application developers.
    This resulted in increased performance when multiple connections are created, decreased code complexity and
    increased compile time safety.
\end{abstract}

\chapter*{Acknowledgements}
I would like to thank my supervisor Dr Colin Perkins for providing this interesting project and providing invaluable
helps and guidence throughout this project.

%==============================================================================

% EDUCATION REUSE CONSENT FORM
% If you consent to your project being shown to future students for educational purposes
% then insert your name and the date below to  sign the education use form that appears in the front of the document. 
% You must explicitly give consent if you wish to do so.
% If you sign, your project may be included in the Hall of Fame if it scores particularly highly.
%
% Please note that you are under no obligation to sign 
% this declaration, but doing so would help future students.
%
\newcommand{\consentname}{Stuart Reilly} % your full name
\newcommand{\consentdate}{\today} % the date you agree
%
\educationalconsent


%==============================================================================
\tableofcontents

%==============================================================================
%% Notes on formatting
%==============================================================================
% The first page, abstract and table of contents are numbered using Roman numerals and are not
% included in the page count. 
%
% From now on pages are numbered
% using Arabic numerals. Therefore, immediately after the first call to \chapter we need the call
% \pagenumbering{arabic} and this should be called once only in the document. 
%
% Do not alter the bibliography style.
%
% The first Chapter should then be on page 1. You are allowed 40 pages for a 40 credit project and 30 pages for a 
% 20 credit report. This includes everything numbered in Arabic numerals (excluding front matter) up
% to but excluding the appendices and bibliography.
%
% You must not alter text size (it is currently 10pt) or alter margins or spacing.
%
%
%==================================================================================================================================
%
% IMPORTANT
% The chapter headings here are **suggestions**. You don't have to follow this model if
% it doesn't fit your project. Every project should have an introduction and conclusion,
% however. 
%
%==================================================================================================================================
\chapter{Introduction}\label{ch:introduction}

% reset page numbering. Don't remove this!
\pagenumbering{arabic}

The way programs have interacted with networks has not changed since the 1980s with the creation of the Berkeley Sockets
API (BSD Sockets).
Since then, programs have seen a drastic increase in relience and range of usage of networks, but they all continue to
use the same BSD Sockets API .
BSD Sockets have become less applicable as the standard API over the years, but no API has been able to replace it.

BSD Sockets was designed to be a simple API which followed the same structure as local file access APIs.
While this approach likely contributed to the APIs widespread use and longevity, it does not match the true nature of
networks.
File access is a synchronous and reliable operation with limited customisation options for how to carry out the
operation.
Networking is inherently asynchronous and unreliable in its operation, and operates on a customisable protocol stack.
While BSD Sockets has been extended over the years to support asynchronous operations with its non-blocking mode, it is
error prone and cumbersome to use.
BSD Sockets does expose a mechanism to customise the protocol stack, but it requires the user of the API to select which
protocols to use from a limited range, supported by the operating system.

The aim of this project is to design and implement a new networking API which overcomes the limitations of the BSD
Sockets API .
The API will be exclusively asynchronous and raise the level of the networking stack the user works with.
It will expose a mechanism to decouple the user from the underlying protocol stack, while allowing the user to specify
constraints on the possible protocols which can be used.

Three major attempts to create a modern alternative to BSD Sockets have been attempted: Post Sockets, NEAT and TAPS.
Post Sockets is an API which aims to decouple programs from the transport-layer protocols
(See Section~\ref{sec:post-sockets}).
NEAT is an API which aims to allow transport-layer protocols to be implemented in user-space
(See Section~\ref{sec:neat}).
TAPS is an API which aims to provide a unified API for connections which communicate the same data conceptually
(See Section~\ref{sec:taps}).

The structure of this document is as follows.
First, a discussion of Post Sockets, NEAT and TAPS as the major influences of this work, and the Rust programming
language is dicussed in Chapter~\ref{ch:background}.
Then in Chapter~\ref{ch:analysis/requirements}, the requirements for the projects are outlined and discussed.
Chapter~\ref{ch:design} then discusses the high-level design of the API the project will create.
Following on from that is a discussion of the implementation of the API in Chapter~\ref{ch:implementation}.
With the implementation in place, Chapter~\ref{ch:evaluation} evaluates the design and the implementation of the API .
Finally, Chapter~\ref{ch:conclusion} discusses possible future work on the project and concludes the document.

%Why should the reader care about what are you doing and what are you actually doing?
%\section{Guidance}
%
%\textbf{Motivate} first, then state the general problem clearly.
%
%\section{Writing guidance}
%\subsection{Who is the reader?}
%
%This is the key question for any writing. Your reader:
%
%\begin{itemize}
%    \item
%    is a trained computer scientist: \emph{don't explain basics}.
%    \item
%    has limited time: \emph{keep on topic}.
%    \item
%    has no idea why anyone would want to do this: \emph{motivate clearly}
%    \item
%    might not know \emph{anything} about your project in particular:
%    \emph{explain your project}.
%    \item
%    but might know precise details and check them: \emph{be precise and
%    strive for accuracy.}
%    \item
%    doesn't know or care about you: \emph{personal discussions are
%    irrelevant}.
%\end{itemize}
%
%Remember, you will be marked by your supervisor and one or more members
%of staff. You might also have your project read by a prize-awarding
%committee or possibly a future employer. Bear that in mind.

%\subsection{References and style guides}
%There are many style guides on good English writing. You don't need to
%read these, but they will improve how you write.

%\begin{itemize}
    %\item
    %\emph{How to write a great research paper} \cite{Pey17} (\textbf{recommended}, even though you aren't writing a research paper)
    %\item
    %\emph{How to Write with Style} \cite{Von80}. Short and easy to read. Available online.
    %\item
    %\emph{Style: The Basics of Clarity and Grace} \cite{Wil09} A very popular modern English style guide.
    %\item
    %\emph{Politics and the English Language} \cite{Orw68}  A famous essay on effective, clear writing in English.
    %\item
    %\emph{The Elements of Style} \cite{StrWhi07} Outdated, and American, but a classic.
    %\item
    %\emph{The Sense of Style} \cite{Pin15} Excellent, though quite in-depth.
%\end{itemize}

%\subsubsection{Citation styles}

%\begin{itemize}
%\item If you are referring to a reference as a noun, then cite it as: ``\citet{Orw68} discusses the role of language in political thought.''
%\item If you are referring implicitly to references, use: ``There are many good books on writing \citep{Orw68, Wil09, Pin15}.''
%\end{itemize}

%There is a complete guide on good citation practice by Peter Coxhead available here: \url{http://www.cs.bham.ac.uk/~pxc/refs/index.html}.
%If you are unsure about how to cite online sources, please see this guide: \url{https://student.unsw.edu.au/how-do-i-cite-electronic-sources}.
%
%\subsection{Plagiarism warning}
%
%\begin{highlight_title}{WARNING}
%
%    If you include material from other sources without full and correct attribution, you are commiting plagiarism. The penalties for plagiarism are severe.
%    Quote any included text and cite it correctly. Cite all images, figures, etc. clearly in the caption of the figure.
%\end{highlight_title}
%

%==================================================================================================================================
\chapter{Background}\label{ch:background}
\input{background}

\chapter{Analysis/Requirements}\label{ch:analysis/requirements}
\input{analysis_requirements}

\chapter{Design}\label{ch:design}
\input{design}

\chapter{Implementation}\label{ch:implementation}
\input{implementation}

\chapter{Evaluation}\label{ch:evaluation}
\input{evaluation}

%==================================================================================================================================
\chapter{Conclusion}\label{ch:conclusion}
The aim of this project was to create a modern alternative to the BSD Sockets API .
BSD Sockets was defined in the 1980s, and has continued to be the standard networking API, despite it not being fit for
modern networked programs.
It fails to represent the true nature of networks and tightly couples programs with the transport protocols used.
This lead to the creation three major attempts to replace BSD Sockets: Post Sockets, NEAT and TAPS.
Post Sockets was designed to decouple programs from the transport-layer protocols.
NEAT was designed to allow transport-layer protocols to be implemented in user-space.
TAPS was designed to provide a unified API for connections which communicate the same data conceptually.
The design of the API took inspiration from the core ideas presented by these three attempts.
This resulted in an API which provides a unified API which decouples programs from transport protocols and provides
support for user-space transport protocols, implemented in Rust, a modern systems programming language.

The resulting API presented a model which follows a user's high-level understanding of a connection.
A connection's configuration is specified first, which can be used to either open a new connection or bind a listener
which listens for connects which match the configuration.
The configuration allows the user to specify the local and remote machines the resulting connection will be betweeen,
and the properties of the connection, such as reliability.
With the connection, the user can then send and receive strongly-typed messages to and from the remote machine.

Two programs written using the API and BSD Sockets results in at least an order of magnitude more code for the BSD
Sockets based program over the program using the API .
Since In general shorter programs are easier to understand, which compounded with increased amount of error handling
code to for the BSD Sockets lead to the program written in the API being easier to understand.
Furthermore, the API is designed as a higher-level library than BSD Sockets, therefore large amounts of the code in the
BSD Sockets program was implemented with the API, such as the DNS lookup.

When the time to establish a connection for each API was compared, BSD Sockets was faster with most domains tested.
The difference was minimal, but was present and was likely due to BSD Sockets being a synchronous API .
Since the API requires an asynchronous runtime to poll the \texttt{Future}s created by the API, the initialisation
of the API was longer than BSD Sockets.
When the same collection of domains was run with similar programs which opened multiple connections for each domain,
the API was significantly faster than BSD Sockets.
Since BSD Sockets is a synchronous API, each connection attempt was done seqentially, leading to an overall slower
program.

\section{Future Work}\label{sec:future-work}
There are four main areas of future work: supporting more properties of transport protocols, include multiple protocol
stacks in the connection racing, implementing more framers and splitting the API crate into multiple smaller crates.

The API only ensures the underlying transport protocol upholds the “reliability'' property.
Other properties defined by TAPS include: preservation of message boundaries and parallel use of multiple paths.
Supporting these properties would allow for more customisation of the protocols used and provide more possible
protocols stacks to include in connection races.

The API only races the IP addresses from a DNS lookup, whereas the TAPS specification specifies multiple protocol
stacks should also be included in the races.
Possible future work would involve including more transport protocols from the operating system, and including them in
the protocol racing.
This would allow for more protocols to be races, increasing their uptake, as well as possible performance improvements.

HTTP is the only framer implemented in the API .
Possible future would involve implementing more application layer protocols such as FTP and NTP .
Since the API does not require these protocols to be implemented within the API, new application layer protocols would
be implemented as separate crates to be distributed separately and the existing HTTP framer would be moved into its
own crate.

The Tokio-based implementation of the API is within the API crate.
This should be moved into its own crate, and the API distributed without an implementation in its crate.
Distributions of the API crate could then include their own default implementation, which can be specialised for the
target platform.

%Summarise the whole project for a lazy reader who didn't read the rest (e.g. a prize-awarding committee).
%\section{Guidance}
%\begin{itemize}
%    \item
%        Summarise briefly and fairly.
%    \item
%        You should be addressing the general problem you introduced in the
%        Introduction.
%    \item
%        Include summary of concrete results (``the new compiler ran 2x
%        faster'')
%    \item
%        Indicate what future work could be done, but remember: \textbf{you
%        won't get credit for things you haven't done}.
%\end{itemize}

%==================================================================================================================================
%
%
%==================================================================================================================================
%  APPENDICES

\begin{appendices}
    \chapter{BSD Sockets Connection Evaluation Code}\label{ch:bsd-sockets-connection-code}
    \lstinputlisting[language=C]{../sockets/main.c}

    \chapter{TAPS Connection Evaluation Code}\label{ch:taps-connection-evaluation-code}
    \lstinputlisting{../taps/src/main.rs}

    \chapter{Bash Script for Latency Testing}\label{ch:bash-script-for-latency-testing}
    \lstinputlisting[language=bash]{../data/test.sh}

    \chapter{Structure of the Source Code}\label{ch:structure-of-the-source-code}
    \dirtree{%
        .1 /.
        .2 connection.rs\DTcomment{Definition of Connection trait}.
        .2 error.rs\DTcomment{Definition of TAPS errors}.
        .2 implementation.rs\DTcomment{Definition of Implementation trait}.
        .2 lib.rs.
        .2 listener.rs\DTcomment{Definition of Listener trait}.
        .2 preconnection.rs\DTcomment{Definition of the Preconnection struct}.
        .2 resolve.rs\DTcomment{Definition of the Endpoint trait}.
        .2 codec\DTcomment{Encode, Decode and Framer traits and implementations}.
        .3 mod.rs.
        .3 decode.rs\DTcomment{Definition of and implementations of Decode trait}.
        .3 encode.rs\DTcomment{Definition of and implementations of Encode trait}.
        .3 error.rs\DTcomment{Definition of decode and deframe error type}.
        .3 frame.rs\DTcomment{Definition of Framer trait}.
        .2 properties.
        .3 mod.rs\DTcomment{Definition of TransportProperties}.
        .3 selection.rs\DTcomment{Definition of SelectionProperties}.
        .2 tokio\DTcomment{Tokio implementation of traits}.
        .3 mod.rs\DTcomment{Implementation of the Implementation trait}.
        .3 connection.rs\DTcomment{Implementation of the Connection trait}.
        .3 error.rs\DTcomment{Definition of Tokio specific errors}.
        .3 race.rs\DTcomment{Connection racing}.
        .3 listener.rs\DTcomment{Implementations of the Listener trait}.
        .4 mod.rs.
        .4 tcp.rs\DTcomment{TCP implementation of the Listener trait}.
    }

%Typical inclusions in the appendices are:
%
%\begin{itemize}
%\item
%  Copies of ethics approvals (required if obtained)
%\item
%  Copies of questionnaires etc. used to gather data from subjects.
%\item
%  Extensive tables or figures that are too bulky to fit in the main body of
%  the report, particularly ones that are repetitive and summarised in the body.
%
%\item Outline of the source code (e.g. directory structure), or other architecture documentation like class diagrams.
%
%\item User manuals, and any guides to starting/running the software.
%
%\end{itemize}
%
%\textbf{Don't include your source code in the appendices}. It will be
%submitted separately.

\end{appendices}

%==================================================================================================================================
%   BIBLIOGRAPHY

% The bibliography style is abbrvnat
% The bibliography always appears last, after the appendices.

\bibliographystyle{abbrvnat}

\bibliography{l4proj}

\end{document}
